---
title: "HadleyFigueroa_R Markdown file"
author: "Hadley Figueroa"
date: "2024-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data description
*	Variable 1: Rank -	Professor’s rank of either assistant professor, associate professor, or professor,
*	Variable 2: Discipline	-	Type of department the professor works in, either applied or theoretical,
*	Variable 3: Years Since PhD	-	Number of years since the professor has obtained their PhD,
*	Variable 4: Years Service	-	Number of years the professor has served the department and/or university,
*	Variable 5: Sex	-	Professor’s sex of either male or female,
*	Variable 6: Salary -	Professor’s nine-month salary (USD)




# Data wrangling
* Step 1. Load the library
* Step 2. Set the working directory
* Step 3. Read csv file into R using the function read_csv(). / loading dataset
* Step 4. Use select(), filter(), group_by(), count() in this part and combine them with pipe operator
* Step 5. Using spread() to transfer the "sex" column to many more columns based on the values in sex.
* Step 6. Reshape the table from wide to long: using gather() to combine multiple "sex" related columns to one column sex

```{r, results=FALSE, message=FALSE}

# 1. Load the library
library(tidyverse)


# 2. Set the working directory
getwd()

# 3. Read csv file into R using the function read_csv(). / loading dataset

Salaries <- read_csv("Salaries.csv")


# 4. Use select(), filter(), group_by(), count() in this part and combine them with pipe operator

## + Assign the output to a new object named Salaries_new
Salaries_new <- Salaries %>%
###   - Select some columns
  select(rank, discipline, sex, salary) %>%
###   - Filter the data where salary is not missing value.
  filter(!is.na(salary)) %>%
###   - Group the data based on the columns of rank, discipline, sex 
  group_by(rank, discipline, sex) %>%
###   - Get the tally count of the observation
  tally()
  



# 5. using spread() to transfer the "sex" column to many more columns based on the values in sex.

# Use spread(): spread(data, key, value, fill = NA, convert = FALSE, drop = TRUE, sep = NULL)
# + Please use the data frame "Salaries_new" for the spread() operation
# + Please split the column sex in the data frame "Salaries_new" to many more columns based on the values in sex
# + Have the values of n(tally count) for the new columns.
# + Fill the missing value with "FALSE" by setting "fill=FALSE"
# + Assign the output to an object "wide"
# + Use pipe operator in this case

wide <- Salaries_new %>%
  spread(key = sex, value= n, fill = FALSE)


# 6. Reshape the table from wide to long: using gather() to combine multiple "sex" related columns to one column sex

# Use gather(): gather(data, key, value, ..., na.rm = FALSE, convert = FALSE, factor_key = FALSE)
# + Please reshape the data frame "wide" obtained from spread() into the long table 
# + In the spread() step, you will get two columns about sex "Female" and "Male"
# + Please make the two sex columns into one column and name the column as "sex"
# + Please have the values in the two sex columns for a new column "n" (tally count)
# + Remove the NA by setting "na.rm=TRUE"
# + Use pipe operator in this case

long <- wide %>%
  gather(key = sex, value = n, Female:Male, na.rm = TRUE)

```